Symbolic linearization of equations of motion of constrained multibody systems

### 2013-08-02 PetersonGedeHubbard2012.tex
Wordsmith section on Kane's method.

\begin{abstract}
% Context
Many common dynamic mechanical systems have configuration or velocity constraints.
% Need  (what we have and what we want)
Analyses of such systems often require linearized forms of the motion
equations.
% Task
To address this issue, we developed a procedure for organizing the constraint
and motion equations and their subsequent linearization.  This procedure was
specifically developed for equations of motion generated by Kane's method, but
is compatible with any method as long as the following are true: only ordinary
differential equations are required to describe the system, and independent and
dependent states remain in the final equations (i.e. no DAEs, no substitution
of variables).
% Object of the document
Following a brief review of Kane's method and the structure of the equations it
generates, we present the procedure to symbolically linearize the nonlinear
motion equations of constrained multibody systems, and illustrate it with the
familiar example of the rolling disc.

@@

%In summary, the procedure is:
%\begin{enumerate}
%    \item Describe the system in the form of the equations shown in Table
%        \ref{table:assumptions}.
%    \item Determine a point of linearization ($\bm{q}^*$, $\dot{\bm{q}}^*$,
%      $\bm{u}^*$, $\dot{\bm{u}}^*$, and $\bm{r}^*$) which satisfies the
%      equations in Table \ref{table:assumptions}.
%    \item Compute the quantities in equation (\ref{eq:quant_to_compute}).
%    \item Identify independent coordinates and speeds; form equations (\ref{eq:C_0}), (\ref{eq:C_1}), and
%        (\ref{eq:C_2}).
%    \item Form equation (\ref{eq:state_space_constrained}), and if needed,
%      equations (\ref{eq:A_prime})-(\ref{eq:B}).
%\end{enumerate}

@@

For a system system with constraints, Kane's dynamical equations\cite{Kane195}
are first ordinary differential equations (in the generalized speeds) which
include both the independent and dependent states (assuming the analyst has not
taken steps to algebraically eliminate states deemed to be dependent). The
dynamical equations and the kinematical differential equations can be solved
for the time derivatives of the generalized coordinates and generalized speeds,
respectively. Taken together, it is tempting to think that this system of first
order differential equations could be linearized in the same fashion as an
unconstrained system. Unfortunately, this is not the case. Linearizing these
equations in the same manner as one would for an unconstrained system (by
directly computing the Jacobian) is incorrect and will yield incorrect
linearized equations (and subsequently incorrect eigenvalues, in the case of a
stability analysis). We illustrate this by means of a familiar example, the
rolling disc, formulated with a non-traditional set of coordinates and
generalized speeds which are not all independent. We then provide a systematic
method to correctly generate the linear equations of motion for constrained
systems which have been found using Kane's method (or any other method which
results in equations with the same form). We then apply this method to the
rolling disc example to illustrate that it yields eigenvalues that match
previously publish results.

We consider systems subject to three types of constraints: configuration,
velocity, and acceleration.  Configuration constraints limit the location or
orientation of parts of the system, relative to the external world or other
parts of the system.  Velocity constraints limit the speeds at which the
configuration can change, either from configuration constraints (through time
differentiation) or independent application.  Velocity constraints most often
appear in systems where there is rolling without slip or there are closed
kinematic loops.  In the context of this paper, acceleration constraints arise
either from time differentiated velocity constraints or similar kinematic
considerations.

% TODO: Improve literature review.
Bottema appears to be the first author to have shown that special
considerations are needed for linearizing nonholonomic systems
\cite{Bottema1949}. While other authors have examined linearization of
nonholonomic systems, we have found that most are not directly applicable to
equations formed using Kane's method, and while we do not doubt that they
\textit{could} be applied correctly, the details of how to do so, as far as the
authors are aware, have not been explicitly presented. Kang et al.
\cite{Kang2003} and Negrut and Ortiz \cite{Negrut2006} have both explored
linearization, but these methods have been developed for equations of motion
found using Lagrange's method, which contain Lagrange multipliers (which are
not present in equations of motion generated from Kane's method). Further, both
of these papers lacked a thorough discussion of basic concepts such as how many
quantities are independent, and in the case of eigenvalue computation, how many
eigenvalues should be present. Minaker and Rieveley \cite{Minaker2010} and
Schwab and Meijaard \cite{Schwab2003} both have developed techniques for
generating equations of motion (and linear forms thereof), however both
techniques are numerical in nature - they do not provide symbolic equations of
motion or symbolic linearized equations. While Neuman presents a symbolic (as
opposed to numeric) formulation for linearization of the Q-matrix formulation
of the Lagrangian\cite{Neuman1984}, his work makes no mention of constrained
systems and attempts to contact the author regarding the software (algebraic
robot modeler) as well as internet searches for the software have been
fruitless.

% TODO:reiterate limitations of model
The goal of this paper is to definitively establish how the first time
derivative of all coordinates and speeds (independent and dependent) depend, to
first order, upon a selection of independent coordinates and independent
speeds, for an arbitrary point of linearization. We demonstrate how a reduced
portion of these linearized equations can be used to perform standard stability
analysis (i.e., for control system design or linear stability analysis). The
formalism presented here is generic enough to cover most examples of
time-varying constrained multibody systems with arbitrary external inputs and
arbitrary specified quantities.
% TODO: electronic supplementary material with examples for other methods
% TODO: rewrite following sentences after e.s.m.
The procedure was created for systems whose equations of motion have been
derived with Kane's method. Although also applicable to dynamic system
equations formulated using other methods, it is restricted to systems which can
be completely described by a set of ODEs (i.e., DAEs needn't be solved).

@@

The benefit of choosing generalized speeds to be other than simply the
derivatives of the corresponding generalized coordinates can be seen in the
following example.
If the orientation of a rigid body were to be described using using Euler
parameters as coordinates, its angular velocity is relatively complicated when
using the time derivatives of the coordinates and requires a velocity
constraint.
Using generalized speeds allows for the angular velocity to be written in the
much simpler form
\begin{align}
\label{eq:ex1}
{^N}\bar{\omega}^B = \it{\omega}_1 \hat{b}_1 + \it{\omega}_2 \hat{b}_2 +
\it{\omega}_3 \hat{b}_3
\end{align}
or, the angular velocity of rigid body $B$ in the reference frame $N$ is the
defined as the sum of three generalized-speed/basis-vector products.
The derivatives of these generalized speeds will then appear in the equations
of motion, rather than the twice time differentiated generalized coordinates
($\ddot{\mathbf{q}}$'s).

A drawback of allowing such definitions is clear when formulating the kinematic
differential equations, which relate the rate of change of the generalized
coordinates to the generalized speeds.
These equations become much more complicated, but this is usually offset by the
significantly simpler equations of motion.

Use and understanding of Kane's method requires another concept, partial
velocities.
Partial velocities are defined as the partial derivatives of the velocity
vectors with respect to the generalized speeds and are defined for both
translational and rotational velocities.
Using the angular velocity of body $B$ in reference frame $N$ as shown in
(\ref{eq:ex1}), the three partial velocities can be written
\[
{^N}\bar{\omega}^B_{\omega_1} = \hat{b}_1 \quad \quad
{^N}\bar{\omega}^B_{\omega_2} = \hat{b}_2 \quad \quad
{^N}\bar{\omega}^B_{\omega_3} = \hat{b}_3
\]
and it can be seen that the partial velocities represent the direction of
motion associated with each generalized speed.

By taking the dot product of each partial velocity and both sides of Newton's
second law, $\bar{F}=m\bar{a}$ (or the rotational equivalent), only terms that
are parallel to the partial velocities remain in the equation.
One equation is generated for each generalized speed.

An important byproduct of this choice is that internal constraint forces
(non-contributing forces) imposed by objects such as pin or sliding joints
no longer need to be considered.

@@

The mathematical formalism for applying Kane's method to multibody systems
follows.
Consider a system composed of rigid bodies $B_1,...,B_g$, particles
$P_1,...,P_h$, with points of force application $Q_1,...,Q_k$, and reference
frames of torque application $E_1,...,E_c$, all defined relative to the
inertial reference frame $N$.
This system has generalized coordinates $q_1,...,q_n$ and generalized speeds
$u_1,...,u_o$.
Additionally, there are $l$ configuration constraints and $m$ velocity
constraints.

In order to properly apply Kane's method, the velocities of each component in
the system need to be written as linear combinations of the generalized speeds.

Translational velocity of particles:
\begin{align}
\label{eq:particle_translational}
{^N}\bar{v}^{P_i} = \sum^o_{j=1} {^N}\bar{v}^{P_i}_{u_j} u_j + {^N}\bar{v}^{P_i}_t
\quad \quad i=1,...,g
\end{align}
Translational velocity of rigid body mass centers:
\begin{align}
\label{eq:rb_translational}
{^N}\bar{v}^{B_i^*} = \sum_{j=1}^o {^N}\bar{v}^{B_i^*}_{u_j} u_j +
{^N}\bar{v}^{B_i^*}_t \quad \quad i=1,...,h
\end{align}
Translational velocity of points of interest:
\begin{align}
\label{eq:points_translational}
{^N}\bar{v}^{Q_i} = \sum_{j=1}^o {^N}\bar{v}^{Q_i}_{u_j} u_j + {^N}\bar{v}^{Q_i}_t
\quad \quad i=1,...,k
\end{align}
Rotational velocity of rigid bodies:
\begin{align}
\label{eq:rb_rotational_bodies}
{^N}\bar{\omega}^{B_i} = \sum_{j=1}^o {^N}\bar{\omega}^{B_i}_{u_j} u_j +
{^N}\bar{\omega}^{B_i}_t \quad \quad i=1,...,h
\end{align}
Rotational velocity of reference frames of interest:
\begin{align}
\label{eq:rb_rotational_frames}
{^N}\bar{\omega}^{E_i} = \sum_{j=1}^o {^N}\bar{\omega}^{E_i}_{u_j} u_j +
{^N}\bar{\omega}^{E_i}_t \quad \quad i=1,...,c
\end{align}

In these equations, each term $\bar{v}_t$ or $\bar{\omega}_t$ represents a
velocity or angular velocity component which is a prescribed function of time,
and has no dependence on generalized speeds (e.g., a crank
driven at a specified rate).

For this system, the configuration constraints are defined as:
\begin{align}
\label{eq:configuration_constraints}
\mathbf{f}_c(\mathbf{q}, t) = \mathbf{0}
\end{align}
where the terms $\mathbf{f}_c$ and $\mathbf{0}$ have the shape $l \times 1$.
The velocity constraints are defined as:
\begin{align}
\label{eq:velocity_constraints}
\mathbf{f}_v(\mathbf{q}, \mathbf{u}, t) = \mathbf{0}
\end{align}
where the terms $\mathbf{f}_v$ and $\mathbf{0}$ have the shape $m \times 1$.
It is important that the velocity constraints include the time-differentiated
configuration constraints or equivalent constraints which produce the same
behavior.
Furthermore, velocity constraints should not be written with $\dot{\mathbf{q}}$
terms (the time derivatives of the coordinates).
Acceleration constraints need to be defined as well.
Typically, they will simply by time-differentiated velocity constraints and
have the form:
\begin{align}
\label{eq:acceleration_constraints}
\mathbf{f}_a(\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, \dot{\mathbf{u}}, t) =
\mathbf{0}
\end{align}

At this point, it is convienient to define a matrix $\mathbf{B}$, which is the
gradient of the velocity constraints $\mathbf{f}_v$ with respect to all of the
generalized speeds $\mathbf{u}$.
\begin{align}
\label{eq:constraint_B}
\mathbf{B} \triangleq \nabla_{\mathbf{u}} \mathbf{f}_v (\mathbf{q}, \mathbf{u},
t)
\end{align}
and from (\ref{eq:velocity_constraints}) and (\ref{eq:constraint_B}) the
following can be written:
\begin{align}
\label{eq:constraint_Bu0}
\mathbf{0} = \mathbf{B}\mathbf{u}
\end{align}

With these definitions established, Kane's method can be applied.
The output of Kane's method, in vector form, is the equation
\begin{align}
\label{eq:kanes_eq}
\mathbf{F} + \mathbf{F}^* = \mathbf{0}
\end{align}
or, the sum of the generalized active forces and generalized inertial forces is
0.
If Newton's second law were written as $f - ma = 0$, the $\mathbf{F}$ is the
multibody generalization of $f$ and $\mathbf{F}^*$ is the multibody
generalization of $-ma$.
These terms are created by taking the dot product of the partial
velocities/angular-velocities, as defined in (\ref{eq:particle_translational})
- (\ref{eq:rb_rotational_frames}), with the resultant forces/torques ($\bar{R}$/
$\bar{T}$) and the inertial forces/torques ($\bar{R}^*$/$\bar{T}^*$).

The resultant force at a point is defined as the sum of all force vectors
acting directly on that point.
The resultant torque on a reference frame is defined similarly - it is the sum
of all torque vectors acting on that reference frame.
Note that distance forces do not need to be transformed into a resultant
force/torque pair for any components of the system; they only need to be
considered at the point of application (e.g., if a force $\bar{f}$ is the only
force applied to body $B$ and is applied at point $Q$, then $\bar{R}_{B^*}=0$
and $\bar{T}_B=0$ but $\bar{R}_Q=\bar{f}$).
These resultant forces and torques are used to write the generalized active
forces:

\begin{equation}
\label{eq:definition_F}
\mathbf{F} =
\begin{bmatrix}
\displaystyle \sum_{i=1}^g \bar{R}_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_1 +
\sum_{i=1}^h \bar{R}_{P_i} \cdot {^N}\bar{v}^{P_i}_1 +
\sum_{i=1}^k \bar{R}_{Q_i} \cdot {^N}\bar{v}^{Q_i}_1 +
\sum_{i=1}^g \bar{T}_{B_i} \cdot {^N}\bar{\omega}^{B_i}_1 +
\sum_{i=1}^c \bar{T}_{E_i} \cdot {^N}\bar{\omega}^{E_i}_1 \\
\displaystyle \sum_{i=1}^g \bar{R}_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_2 +
\sum_{i=1}^h \bar{R}_{P_i} \cdot {^N}\bar{v}^{P_i}_2 +
\sum_{i=1}^k \bar{R}_{Q_i} \cdot {^N}\bar{v}^{Q_i}_2 +
\sum_{i=1}^g \bar{T}_{B_i} \cdot {^N}\bar{\omega}^{B_i}_2 +
\sum_{i=1}^c \bar{T}_{E_i} \cdot {^N}\bar{\omega}^{E_i}_2 \\
\displaystyle \vdots \\
\displaystyle \sum_{i=1}^g \bar{R}_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_o +
\sum_{i=1}^h \bar{R}_{P_i} \cdot {^N}\bar{v}^{P_i}_o +
\sum_{i=1}^k \bar{R}_{Q_i} \cdot {^N}\bar{v}^{Q_i}_o +
\sum_{i=1}^g \bar{T}_{B_i} \cdot {^N}\bar{\omega}^{B_i}_o +
\sum_{i=1}^c \bar{T}_{E_i} \cdot {^N}\bar{\omega}^{E_i}_o
\end{bmatrix}
\end{equation}

The inertia forces and torques, $\bar{R}^*$ and $\bar{T}^*$, will now be
defined. Note: the reference frame these are defined is not written - at this
point in the derivation of the equations of motion, they are assumed to be in
the inertial frame which is $N$ for this system.

Particles:
\begin{align}
\label{eq:particle_gen_inertia}
\bar{R}^*_P \triangleq -m_p {^N}\bar{a}^P
\end{align}
Rigid Bodies:
\begin{align}
\label{eq:rb_translational_gen_inertia}
\bar{R}^*_{B^*} \triangleq -m_B {^N}\bar{a}^{B^*}
\end{align}
\begin{align}
\label{eq:rb_rotational_gen_inertia}
\bar{T}^*_B \triangleq -{^N}\bar{\alpha}^B \cdot \bar{\bar{I}}^{B/B^*} -
{^N}\bar{\omega}^B \times \bar{\bar{I}}^{B/B^*} \cdot {^N}\bar{\omega}^B
\end{align}
Where $\bar{\bar{I}}^{B/B^*}$ is the central inertia dyadic of the rigid body
$B$ about its mass center $B^*$ \cite{Kane1985}.

If the basis vectors of the reference frame associated with the rigid body $B$
are aligned with the principal axes of $B$, then the following definitions can
be used
\begin{align}
\alpha_i \triangleq {^N}\bar{\alpha}^B \cdot \hat{b}_i \quad \quad
\omega_i \triangleq {^N}\bar{\omega}^B \cdot \hat{b}_i \quad \quad
I_i \triangleq \hat{b}_i \cdot \bar{\bar{I}}^{B/B^*} \cdot \hat{b}_i
\quad \quad \text{for }i=1,2,3
\end{align}
In this case, $I_1$, $I_2$, and $I_3$ will be the principal inertias of rigid
body $B$.
This allows the inertia torque to be written
\begin{align}
\label{eq:simple_rb_rotational_gen_inertia}
\bar{T}_B^* =& -[\alpha_1 I_1 - \omega_2 \omega_3 (I_2 - I_3)] \hat{b}_1\\
             & -[\alpha_2 I_2 - \omega_3 \omega_1 (I_3 - I_1)] \hat{b}_2\\
             & -[\alpha_3 I_3 - \omega_1 \omega_2 (I_1 - I_2)] \hat{b}_3\\
\end{align}

Now $\mathbf{F}^*$ can be constructed in a manner similar to $\mathbf{F}$.

>>>>

\begin{abstract}
% Context
Many common dynamic mechanical systems have configuration or velocity constraints.
% Need  (what we have and what we want)
Analyses of such systems often require linearized forms of the motion
equations.
% Task
To address this issue, we developed a procedure for organizing the constraint
and motion equations and their subsequent linearization.  This procedure was
specifically developed for equations of motion generated by Kane's method, but
is compatible with any method as long as the following are true: only ordinary
differential equations are required to describe the system, and independent and
dependent states remain in the final equations (i.e. no DAE's, no substitution
of variables).
% Object of the document
Following a brief review of Kane's method and the structure of the equations it
generates, we present the procedure to symbolically linearize the nonlinear
motion equations of constrained multibody systems, and illustrate it with the
familiar example of the rolling disk.

@@

@@

For a system system with constraints, Kane's dynamical equations\cite{Kane1985}
are first order ODE's (in the generalized speeds) which include both the
independent and dependent states (assuming the dependents states have not been
algebraically eliminated). The dynamical equations and the kinematical
differential equations can be solved for the time derivatives of the
generalized coordinates and generalized speeds, respectively. Taken together,
it is tempting to think that this system of first order differential equations
could be linearized in the same fashion as an unconstrained system.
Unfortunately, this is not the case. Linearizing these equations in the same
manner as one would for an unconstrained system (by directly computing the
Jacobian) is incorrect and will yield incorrect linearized equations (and
subsequently incorrect eigenvalues, in the case of a stability analysis). We
illustrate this by means of a familiar example, the rolling disk, formulated
with a non-traditional set of coordinates and generalized speeds which are not
all independent. We then provide a systematic method to correctly generate the
linear equations of motion for constrained systems which have been found using
Kane's method (or any other method which results in equations with the same
form). We then apply this method to the rolling disk example to illustrate that
it yields eigenvalues that match previously published results.

We consider three types of constraints: configuration, velocity, and
acceleration. Configuration constraints limit the location or orientation of
parts of the system, relative to the external world or other parts of the
system. Velocity constraints limit the speeds at which the configuration can
change, either from configuration constraints (through time differentiation) or
independent application. Velocity constraints most often appear in systems
where there is rolling without slip or there are closed kinematic loops. In the
context of this paper, acceleration constraints arise either from time
differentiated velocity constraints or similar kinematic considerations.

% TODO: Improve literature review.
Bottema appears to be the first author to have shown that special
considerations are needed for linearizing nonholonomic systems
\cite{Bottema1949}. While other authors have examined linearization of
nonholonomic systems, we have found that most techniques are not directly
applicable to equations formed using Kane's method, and while we do not doubt
that they \textit{could} be applied correctly, the details of how to do so, as
far as the authors are aware, have not been explicitly presented. Kang et al.
\cite{Kang2003} and Negrut and Ortiz \cite{Negrut2006} have both explored
linearization, but these methods have been developed for equations of motion
found using Lagrange's method, which contain Lagrange multipliers (which are
not present in equations of motion generated from Kane's method). Further, both
of these papers lack a discussion of basic concepts such as how many quantities
are independent, and in the case of eigenvalue computation, how many
eigenvalues should be present. Minaker and Rieveley \cite{Minaker2010} and
Schwab and Meijaard \cite{Schwab2003} both have developed techniques for
generating equations of motion (and linear forms thereof), however both
techniques are numerical in nature -- they do not provide symbolic equations of
motion or symbolic linearized equations. While Neuman presents a symbolic (as
opposed to numeric) formulation for linearization of the Q-matrix formulation
of the Lagrangian\cite{Neuman1984}, his work makes no mention of constrained
systems and attempts to contact the author regarding the software (algebraic
robot modeler -- Arm) as well as internet searches for the software have been
fruitless.

% TODO:reiterate limitations of model
The goal of this paper is to definitively establish how the first time
derivative of all coordinates and speeds (independent \textit{and} dependent)
depend, to first order, upon a selection of independent coordinates and
independent speeds, for an arbitrary point of linearization. We demonstrate how
a reduced portion of these linearized equations can be used to perform standard
stability analysis (i.e., for control system design or linear stability
analysis). The formalism presented here is generic enough to cover most
examples of time-varying constrained multibody systems with arbitrary external
inputs and arbitrary specified quantities.
% TODO: electronic supplementary material with examples for other methods
% TODO: rewrite following sentences after e.s.m.
The procedure was created for systems whose equations of motion have been
derived with Kane's method. Although also applicable to dynamic system
equations formulated using other methods, it is restricted to systems which can
be completely described by a set of ODE's (i.e., DAE's needn't be solved).

@@

The benefit of choosing generalized speeds to be other than simply the
derivatives of the corresponding generalized coordinates can be seen in the
following example. If the orientation of a rigid body were to be described
using using Euler parameters or quaternions as coordinates, its angular
velocity in terms of time derivatives of those coordinates is relatively
complicated; it also requires a velocity constraint. Using generalized speeds
allows for the angular velocity to be expressed more simply as
\begin{align}
\label{eq:ex1}
{^N}\bar{\omega}^B = \it{\omega}_1 \hat{b}_1 + \it{\omega}_2 \hat{b}_2 +
\it{\omega}_3 \hat{b}_3
\end{align}
or, the angular velocity of rigid body $B$ in the reference frame $N$ is the
defined as the sum of three generalized-speed/basis-vector products.
The derivatives of these generalized speeds will then appear in the equations
of motion, rather than the twice time differentiated generalized coordinates
($\ddot{\mathbf{q}}$'s).

A drawback of allowing such definitions can be seen when formulating the
kinematic differential equations, which relate the rate of change of the
generalized coordinates to the generalized speeds. These equations become more
complicated (in comparison to $\dot{q}_i = u_i$), but this is usually offset by
the significantly simpler dynamical equations of motion\cite{Mitiguy1996}

Use and understanding of Kane's method requires the use and understanding of
partial velocities and partial angular velocities. These are simply the the
partial derivatives of the (angular) velocity vectors with respect to the
generalized speeds. For example, for the angular velocity in equation
(\ref{eq:ex1}), the partial angular velocities of body $B$ in reference frame
$N$ are
\[
{^N}\bar{\omega}^B_{\omega_1} = \hat{b}_1 \quad \quad
{^N}\bar{\omega}^B_{\omega_2} = \hat{b}_2 \quad \quad
{^N}\bar{\omega}^B_{\omega_3} = \hat{b}_3
\]
and it can be seen that the partial velocities represent the direction of
motion associated with each generalized speed. The geometric interpretations of
these partial (angular) velocities are discussed at length by
Lesser\cite{Lesser1992}.

Taking the dot product of each partial velocity and both sides of Newton's
second law, $\bar{F}=m\bar{a}$ (or the rotational equivalent), only terms that
are parallel to the partial velocities remain; one scalar equation is generated
for each generalized speed. An important byproduct is that internal constraint forces
(non-contributing forces) imposed by objects such as pin or sliding joints no
longer need to be considered. The reason for this is that partial velocities
are \textit{by construction} perpendicular to forces of constraint. Constraint
forces can be included in the formulation, but will not appear in the final
equations of motion because of this very desirable property.

@@

The mathematical formalism for applying Kane's method to multibody systems
follows. Consider a system composed of particles $P_1,...,P_h$, rigid bodies
$B_1,...,B_g$, with points of force application $Q_1,...,Q_k$, and reference
frames of torque application $E_1,...,E_c$, all defined relative to the
inertial reference frame $N$. This system has generalized coordinates
$q_1,...,q_n$ and generalized speeds $u_1,...,u_o$. Additionally, there are $l$
configuration constraints and $m$ velocity constraints.

In order to properly apply Kane's method, the velocities of each component in
the system need to be written as linear combinations of the generalized speeds.
The translational velocity of the particles are
\begin{align}
\label{eq:particle_translational}
{^N}\bar{v}^{P_i} = \sum^o_{j=1} {^N}\bar{v}^{P_i}_{u_j} u_j + {^N}\bar{v}^{P_i}_t
\quad \quad i=1,...,g
\end{align}
while the translational velocity of rigid body mass centers are
\begin{align}
\label{eq:rb_translational}
{^N}\bar{v}^{B_i^*} = \sum_{j=1}^o {^N}\bar{v}^{B_i^*}_{u_j} u_j +
{^N}\bar{v}^{B_i^*}_t \quad \quad i=1,...,h
\end{align}
and finally the translational velocity of points which have applied forces are
\begin{align}
\label{eq:points_translational}
{^N}\bar{v}^{Q_i} = \sum_{j=1}^o {^N}\bar{v}^{Q_i}_{u_j} u_j + {^N}\bar{v}^{Q_i}_t
\quad \quad i=1,...,k
\end{align}
Next, the angular velocity of rigid bodies are
\begin{align}
\label{eq:rb_rotational_bodies}
{^N}\bar{\omega}^{B_i} = \sum_{j=1}^o {^N}\bar{\omega}^{B_i}_{u_j} u_j +
{^N}\bar{\omega}^{B_i}_t \quad \quad i=1,...,h
\end{align}
and the angular velocity of reference frames which have applied torques are
\begin{align}
\label{eq:rb_rotational_frames}
{^N}\bar{\omega}^{E_i} = \sum_{j=1}^o {^N}\bar{\omega}^{E_i}_{u_j} u_j +
{^N}\bar{\omega}^{E_i}_t \quad \quad i=1,...,c
\end{align}

In these equations, each term $\bar{v}_t$ or $\bar{\omega}_t$ represents a
velocity or angular velocity component which is a prescribed function of time,
and has no dependence on generalized speeds (e.g., a motor or crank driven at a
specified rate).

For this system, we consider configuration constraints of the form
\begin{align}
\label{eq:configuration_constraints}
\mathbf{f}_c(\mathbf{q}, t) = \mathbf{0}
\end{align}
where the terms $\mathbf{f}_c$ and $\mathbf{0}$ have the shape $l \times 1$.
The velocity constraints are taken to have the form
\begin{align}
\label{eq:velocity_constraints}
\mathbf{f}_v(\mathbf{q}, \mathbf{u}, t) = \mathbf{0}
\end{align}
where the terms $\mathbf{f}_v$ and $\mathbf{0}$ have the shape $m \times 1$,
and $\mathbf{f}_v$ depends linearly on $\mathbf{u}$.  It is important that the
velocity constraints include the time-differentiated configuration constraints
or equivalent constraints which produce the same behavior. Furthermore,
velocity constraints should be written such that generalized speeds and
coordinates appear, but $\dot{\mathbf{q}}$ terms do not.  Finally, acceleration
constraints are typically time-differentiated velocity constraints (or
similarly formed from kinematic considerations) and are assumed to have the
form
\begin{align}
\label{eq:acceleration_constraints}
\mathbf{f}_a(\mathbf{q}, \dot{\mathbf{q}}, \mathbf{u}, \dot{\mathbf{u}}, t) =
\mathbf{0}
\end{align}
In contrast to the velocity constraints, our formulation permits the
acceleration constraints to involve $\dot{\mathbf{q}}$ terms (althernatively
they can be eliminated by appealing to the kinematic differential equations).

By assumption, the velocity constraints are linear with respect to
$\mathbf{u}$, so we introduce
\begin{align}
\label{eq:constraint_B}
\mathbf{B} &\triangleq \nabla_{\mathbf{u}} \mathbf{f}_v (\mathbf{q}, \mathbf{u},
t)
\end{align}
which implies
\begin{align}
\label{eq:constraint_Bu0}
\mathbf{B}\mathbf{u} + \mathbf{f}_{vt}(\mathbf{q}, t) &= \mathbf{0}
\end{align}
where $\mathbf{f}_{vt}$ arises from explicit time varying terms that occur in
the velocity constraints. These typically arise from specified quantites whose
dynamics are being ignored (e.g., a motor spinning at constant speed).

With these definitions established, Kane's method can be applied. In vector
form, Kane's dynamical equations are
\begin{align}
\label{eq:kanes_eq}
\mathbf{F} + \mathbf{F}^* = \mathbf{0}
\end{align}
where $\mathbf{F}$ are the generalized active forces and $\mathbf{F}^*$ are the
generalized inertia forces. If Newton's second law is written as $f - ma = 0$,
$\mathbf{F}$ is the multibody generalization of $f$ and $\mathbf{F}^*$ is the
multibody generalization of $-ma$. As will be shown, these forces are obtained
by taking the dot product of the partial velocities/angular-velocities, as
defined in (\ref{eq:particle_translational}) - (\ref{eq:rb_rotational_frames}),
with the resultant forces/torques ($\bar{R}$/ $\bar{T}$) and with inertial
forces/torques ($\bar{R}^*$/$\bar{T}^*$).

The resultant force at a point is defined as the sum of all force vectors
acting directly on that point. The resultant torque on a reference frame is
defined similarly -- it is the sum of all torque vectors acting on that
reference frame. Distance forces do not need to be transformed into a resultant
force/torque pair for any components of the system; they only need to be
considered at the point of application (e.g., if a force $\bar{f}$ is the only
force applied to body $B$ at point $Q$, then $\bar{R}_{B^*}=0$ and
$\bar{T}_B=0$ but $\bar{R}_Q=\bar{f}$). The generalized active forces are
\begin{equation}
\label{eq:definition_F}
\mathbf{F} =
\begin{bmatrix}
\displaystyle \sum_{i=1}^g \bar{R}_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_1 +
\sum_{i=1}^h \bar{R}_{P_i} \cdot {^N}\bar{v}^{P_i}_1 +
\sum_{i=1}^k \bar{R}_{Q_i} \cdot {^N}\bar{v}^{Q_i}_1 +
\sum_{i=1}^g \bar{T}_{B_i} \cdot {^N}\bar{\omega}^{B_i}_1 +
\sum_{i=1}^c \bar{T}_{E_i} \cdot {^N}\bar{\omega}^{E_i}_1 \\
\displaystyle \sum_{i=1}^g \bar{R}_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_2 +
\sum_{i=1}^h \bar{R}_{P_i} \cdot {^N}\bar{v}^{P_i}_2 +
\sum_{i=1}^k \bar{R}_{Q_i} \cdot {^N}\bar{v}^{Q_i}_2 +
\sum_{i=1}^g \bar{T}_{B_i} \cdot {^N}\bar{\omega}^{B_i}_2 +
\sum_{i=1}^c \bar{T}_{E_i} \cdot {^N}\bar{\omega}^{E_i}_2 \\
\displaystyle \vdots \\
\displaystyle \sum_{i=1}^g \bar{R}_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_o +
\sum_{i=1}^h \bar{R}_{P_i} \cdot {^N}\bar{v}^{P_i}_o +
\sum_{i=1}^k \bar{R}_{Q_i} \cdot {^N}\bar{v}^{Q_i}_o +
\sum_{i=1}^g \bar{T}_{B_i} \cdot {^N}\bar{\omega}^{B_i}_o +
\sum_{i=1}^c \bar{T}_{E_i} \cdot {^N}\bar{\omega}^{E_i}_o
\end{bmatrix}
\end{equation}

We now define the inertia forces and torques, $\bar{R}^*$ and $\bar{T}^*$. We
leave out the inertial frame $N$ from the expressions; it is assumed to be $N$.
The inertia force for particle $P$ is
\begin{align}
\label{eq:particle_gen_inertia}
\bar{R}^*_P \triangleq -m_P {^N}\bar{a}^P
\end{align}
while the inertia force and inertia torque for rigid body $B$ is
\begin{align}
\label{eq:rb_translational_gen_inertia}
\bar{R}^*_{B^*} &\triangleq -m_B {^N}\bar{a}^{B^*} \\
\label{eq:rb_rotational_gen_inertia}
\bar{T}^*_B &\triangleq -{^N}\bar{\alpha}^B \cdot \bar{\bar{I}}^{B/B^*} -
{^N}\bar{\omega}^B \times \bar{\bar{I}}^{B/B^*} \cdot {^N}\bar{\omega}^B
\end{align}
Where $\bar{\bar{I}}^{B/B^*}$ is the central inertia dyadic of the rigid body
$B$ about its mass center $B^*$ \cite{Kane1985}. The generalized inertia forces
are
\begin{equation}
\label{eq:definition_Fstar}
\mathbf{F}^* =
\begin{bmatrix}
\displaystyle \sum_{i=1}^g \bar{R}^*_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_1 +
\sum_{i=1}^h \bar{R}^*_{P_i} \cdot {^N}\bar{v}^{P_i}_1 +
\sum_{i=1}^g \bar{T}^*_{B_i} \cdot {^N}\bar{\omega}^{B_i}_1 \\
\displaystyle \sum_{i=1}^g \bar{R}^*_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_2 +
\sum_{i=1}^h \bar{R}^*_{P_i} \cdot {^N}\bar{v}^{P_i}_2 +
\sum_{i=1}^g \bar{T}^*_{B_i} \cdot {^N}\bar{\omega}^{B_i}_2 \\
\displaystyle \vdots \\
\displaystyle \sum_{i=1}^g \bar{R}^*_{B^*_i} \cdot {^N}\bar{v}^{B^*_i}_o +
\sum_{i=1}^h \bar{R}^*_{P_i} \cdot {^N}\bar{v}^{P_i}_o +
\sum_{i=1}^g \bar{T}^*_{B_i} \cdot {^N}\bar{\omega}^{B_i}_o
\end{bmatrix}
\end{equation}

All the terms in (\ref{eq:kanes_eq}) are now defined, but this equation is
valid only if there are no velocity constraints. If there are velocity
constraints, the equations must be transformed. As we present them, Kane's
nonholonomic dynamical equations will be of dimension $o - m$, where $o$ is the
total number of generalized speeds and $m$ is the number of velocity
constraints. However, these $o-m$ equations generally involve all $o$ speeds
(we assume the dependent speeds and their time derivatives have \textit{not}
been algebraically eliminated). In order to obtain the nonholonomic dynamical
equations, however, a choice must be made as to which speeds will be dependent.
The methods for doing this are outside the scope of this paper, but it is
important that whatever selection is made avoids singularities which can occur
in certain configurations or sets of system parameters. For a careful treatment
of this, we encourage the reader to examine the work of Reckdahl
\cite{Reckdahl1996}.
